Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-11-19 draft.tex
small typo fixes

The international system is characterized by a norm under which seceding entities are rarely recognized by foreign states unless they are first recognized by the home state.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states. In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm makes sense in a context where almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby reducing the incentives for secessionist movements.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize." For the definitive international legal treatment of recognition, see Crawford (2006).}%The higher the costs of secession, the greater the security of existing states. 
Thus, unrecognized states emerge when seceding entities gain \emph{de facto} territorial control through successful wars of secession, but are not militarily strong enough to force recognition of their independence from the home state from which they have seceded.

@@

We acknowledge that there may exist patrons whose most-preferred outcome is the status quo. This naturally makes a status quo equilibrium easier to achieve, as there is an actor who strictly prefers the status quo and can expend resources to make it more likely. We choose to examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. Even in this circumstance, the status quo remains an equilibrium outcome.

@@

\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, Restriction (5) is more likely to be binding than Restriction (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be binding.}

@@

In equilibrium, the outside actors' investments are paired: the patron's investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding. Likewise, investments by player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding. Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting. In each case, the potential  investments are such that the largest continuation payoffs in the game between the secessionists and the government are those from playing status quo.

@@

To understand the construction of this equilibrium, notice that there are three possible ways to disturb the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order for the Status Quo Equilibrium to persist, each of these three disturbances must be deterred.

Since the patron moves first, in the Status Quo Equilibrium, the only investment that takes place is the patron's investment in the status quo payoffs of the secessionists to deter the international community from provoking the secessionists to cede the issue of sovereignty. The international community's willingness to counteract investments by the patron toward the other two disturbances implies that there will be no investments in equilibrium in cases (2) and (3).

>>>>

The international system is characterized by a norm under which seceding entities are rarely recognized by foreign states unless they are first recognized by the home state.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states. In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm makes sense in a context where almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby reducing the incentives for secessionist movements.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize." For the definitive international legal treatment of recognition, see Crawford (2006).} Thus, unrecognized states emerge when seceding entities gain \emph{de facto} territorial control through successful wars of secession, but are not militarily strong enough to force recognition of their independence from the home state from which they have seceded.

@@

We acknowledge that there may exist patrons whose most-preferred outcome is the status quo. This naturally makes a status quo equilibrium easier to achieve, as there is an actor who strictly prefers the status quo and can expend resources to make it more likely. We choose to examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. We will show that even in this circumstance the status quo remains an equilibrium outcome.

@@

\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}

@@

We now turn to defining and describing the Status Quo Equilibrium. We will see that, in this equilibrium, the outside actors' investments are paired: the patron's investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding. Likewise, investments by player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding. Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting. In each case, the potential  investments are such that the largest continuation payoffs in the game between the secessionists and the government are those from playing status quo.

@@

To understand the construction of this equilibrium, notice that there are three possible ways to disrupt the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order to establish that the Status Quo Equilibrium exists, it must be shown that each of these three deviations will be deterred.

Since the patron moves first, the only investment that takes place in the Status Quo Equilibrium is the patron's investment in the status quo payoffs of the secessionists to deter the international community from provoking the secessionists to cede the issue of sovereignty. The international community's willingness to counteract investments by the patron toward the other two disturbances implies that there will be no investments in equilibrium in cases (2) and (3).
